---
title: "Computational Modeling - Assignment 2 Part 1"
author: "Lasse Damgaard"
date: "4/20/2018"
output: html_document
---

```{r}

setwd("C:/Users/Bruger/Desktop/CogSci/Classes/Computational Modeling/Data")
library(pacman)
p_load(brms, patchwork, gridGraphics, coda, mvtnorm, devtools, rethinking)
```

# You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

# To keep things simple (your teachers should not be faced with too complicated things):
# - You created a pool of equally challenging questions on CogSci
# - Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
# - Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

# This is the data:
# - Riccardo: 3 correct answers out of 6 questions
# - Kristian: 2 correct answers out of 2 questions (then he gets bored)
# - Josh: 160 correct answers out of 198 questions (Josh never gets bored)
# - Mikkel: 66 correct answers out of 132 questions

```{r}

data <- data.frame(
  Correct = c(3,2,160,66),
  Questions = c(6,2,198,132),
  Teacher = c("RF","KT","JS","MW"))

```

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

```{r}

dens <- 100 
p_grid <- seq( from = 0 , to = 1 , length.out = dens ) 
prior <- rep( 1 , dens ) 
likelihood <- dbinom( 6 , size = 9 , prob = p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior) 

temp_df<- data.frame(p_grid = p_grid, posterior = posterior)
ggplot(temp_df, aes(p_grid, posterior))+
  geom_line()+
  labs(x = "Score", y = "Probability", title = "Riccardo's Uninformed Posterior")

```

- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}

cs.qa = map(
  alist(
    c ~ dbinom(6,p), 
    p ~ dunif(0,1) 
    ) ,
  data = list(c = 3)) 

precis(cs.qa)

dummy <- rbinom( 100 , size = 6 , prob = 0.5 )
simplehist( dummy , main = "Riccardo Dummy Data", xlab = "Answers Correct", ylab = "Probability" )

```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.

```{r}

prior <- rep( 1 , dens )
temp_df <- data.frame(p_grid = p_grid, prior = prior)
ggplot(temp_df, aes(p_grid, prior))+
  geom_line()+
  labs(x = "Score", y = "Probability", title = "Uniform Prior")

teachers <- c("RF","KT","JS","MW")
for (teacher in teachers){
  loop_data <- subset(data, Teacher == teacher )
  dens <- 1000 
  p_grid <- seq( from = 0 , to = 1 , length.out = dens ) 
  prior <- rep( 1 , dens ) 
  likelihood <- dbinom(loop_data$Correct , size = loop_data$Questions , prob = p_grid )
  unstd.posterior <- likelihood * prior 
  posterior <- unstd.posterior / sum(unstd.posterior) 


  temp_df <- data.frame(p_grid = p_grid, posterior = posterior)
  temp <- ggplot(temp_df, aes(p_grid, posterior))+
   geom_line()+
   labs(x = "Score", y = "Probability")
   assign(paste(teacher, "Uninformed_Posterior", sep = "_"), temp)
}

```

2a. Produce plots of the prior, and posterior for each teacher.

```{r}

gridExtra::grid.arrange(JS_Uninformed_Posterior , RF_Uninformed_Posterior, KT_Uninformed_Posterior, MW_Uninformed_Posterior )

```

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?

```{r}

dens <- 100
p_grid <- seq( from = 0 , to = 1 , length.out= dens )
prior <- dnorm(p_grid,0.5,0.01)
temp_df<- data.frame(p_grid = p_grid, prior = prior)
ggplot(temp_df, aes(p_grid, prior))+
  geom_line()+
  labs(x = "Score", y = "Probability", title = "Extremely Skeptical Prior")

for (teacher in teachers){
  loop_data <- subset(data, Teacher == teacher )
  dens <- 1000
  p_grid <- seq( from = 0 , to = 1 , length.out = dens )
  prior = dnorm(p_grid,0.8,0.2)
  likelihood <- dbinom(loop_data$Correct , size = loop_data$Questions , prob = p_grid)
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)

  temp_df <- data.frame(p_grid = p_grid, posterior = posterior)
  temp <-ggplot(temp_df, aes(p_grid, posterior))+
   geom_line()
 assign(paste(teacher, "informed", sep = "_"), temp)
}

```

3a. Produce plots of the prior and posterior for each teacher.

```{r}

gridExtra::grid.arrange(JS_informed , RF_informed, KT_informed, MW_informed )

```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}

data$Correct <- data$Correct * 100
data$Questions <- data$Questions * 100

for (teacher in teachers){
  loop_data <- subset(data, Teacher == teacher )
  dens <- 1000
  p_grid <- seq( from = 0 , to = 1 , length.out= dens )
  prior <- rep( 1 , dens)
  likelihood <- dbinom(loop_data$Correct , size = loop_data$Questions , prob = p_grid)
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  temp_df <- data.frame(p_grid = p_grid, posterior = posterior)
  temp <- ggplot(temp_df, aes(p_grid, posterior))+
   geom_line()
 assign(paste(teacher, "uninformed", sep = "_"), temp)
}

gridExtra::grid.arrange(JS_uninformed , RF_uninformed, KT_uninformed, MW_uninformed )

for (teacher in teachers){
  loop_data <- subset(data, Teacher == teacher )
  dens <- 1000
  p_grid <- seq( from = 0 , to = 1 , length.out = dens )
  prior <- dnorm(p_grid,0.8,0.2)
  likelihood <- dbinom(loop_data$Correct , size = loop_data$Questions , prob = p_grid)
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)

  temp_df <- data.frame(p_grid = p_grid, posterior = posterior)
  temp <- ggplot(temp_df, aes(p_grid, posterior))+
   geom_line()+
   labs(x = "Score", y = "Probability")
 assign(paste(teacher, "informed", sep = "_"), temp)
}

gridExtra::grid.arrange(JS_informed , RF_informed, KT_informed, MW_informed )

for (teacher in teachers){ 
  loop_data <- subset(data, Teacher == teacher)
  cs.qa = map(
    alist(
      c ~ dbinom(loop_data$Questions,p), 
      p ~ dunif(0,1) 
    ) ,
    data = list(c = loop_data$Correct)) 
  output <- precis(cs.qa)
  print(output)
  cs.qa = map(
    alist(
      c ~ dbinom(loop_data$Questions,p), 
      p ~ dunif(0.8,0.2) 
    ) ,
    data = list(c = loop_data$Correct)) 
  output <- precis(cs.qa)
  print(output)
}

```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

```{r}

dens <- 1000
p_grid <- seq( from = 0 , to = 1 , length.out = dens )
prior = pnorm(p_grid, mean = 0.5, sd = 0.1, lower.tail = FALSE, log.p = FALSE)

temp_df <- data.frame(p_grid = p_grid, prior = prior)
ggplot(temp_df, aes(p_grid, prior))+
   geom_line()+
   labs(x = "Score", y = "Probability", title = "Skeptical Prior")

```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

```{r}

R_PositiveModel <- brm(Correct|trials(Questions)~1, data = subset(data,Teacher == "RF"), prior = prior("normal(0.8,0.2)", class = "Intercept"), family = 'binomial')

K_PositiveModel <- brm(Correct|trials(Questions)~1, data = subset(data,Teacher == "KT"), prior = prior("normal(0.8,0.2)", class = "Intercept"), family = 'binomial')

J_PositiveModel <- brm(Correct|trials(Questions)~1, data = subset(data,Teacher == "JS"), prior = prior("normal(0.8,0.2)", class = "Intercept"), family = 'binomial')

M_PositiveModel <- brm(Correct|trials(Questions)~1, data = subset(data,Teacher == "MW"), prior = prior("normal(0.8,0.2)", class = "Intercept"), family = 'binomial')

plot(R_PositiveModel)
plot(K_PositiveModel)
plot(J_PositiveModel)
plot(M_PositiveModel)

```

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.

```{r}

FlatModel <- brm(Correct|trials(Questions)~1, data = subset(data,Teacher == "RF"), prior = prior("uniform(0,1)", class = "Intercept"), family = 'binomial')
plot(FlatModel)
temp <- grab_grob()
assign("Flat", temp)

PositiveModel <- brm(Correct|trials(Questions)~1, data=subset(data,Teacher == "RF"), prior = prior("normal(0.8,0.2)", class = "Intercept"), family = 'binomial')
plot(PositiveModel)
temp <- grab_grob()
assign("Positive", temp)

SkepticalModel <- brm(Correct|trials(Questions)~1,data = subset(data,Teacher=="RF"), prior = prior("normal(0.5,0.01)", class = "Intercept"), family = 'binomial')
plot(SkepticalModel)
temp <- grab_grob()
assign("Skeptical", temp)

gridExtra::grid.arrange(FlatModel, PositiveModel, SkepticalModel)

```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?

```{r}

k = map(
  alist(
    c ~ dbinom(2,p), 
    p ~ dunif(0,1) 
  ) ,
  data = list(c = 2)) 

precis(k)


j = map(
  alist(
    c ~ dbinom(198,p), 
    p ~ dunif(0,1) 
  ) ,
  data = list(c = 160)) 

precis(j)

```

END
